Modelling the City of Melbourne Pedestrian Data

### 2017-05-07 presentation/proposal_pres.Rmd
still need to add: shiny app info, limitations, q&a slide, acknowledgements, more visualisation

  melb <- get_map(location=c(mean(range(ped_loc$Longitude)),
                               mean(range(ped_loc$Latitude))),
                                zoom=14)
```

@@

```{r, echo = FALSE}
      ped_df$Year <- as.factor(ped_df$Year)
      ped_df$Month <- as.factor(ped_df$Month)
      ped_df$Month <- factor(ped_df$Month, levels(ped_df$Month)[c(5,4,8,1,9,7,6,2,12,11,10,3)])
      ped_df$Time <- as.factor(ped_df$Time)
      ped_df$Timeplot <- as.integer(ped_df$Time)
      ped_df$MTime <- as.integer(ped_df$Time) + ((as.integer(ped_df$Mdate)-1) * 24)
      ped_df$Day <- as.factor(ped_df$Day)
      ped_df$Day <- factor(ped_df$Day, levels(ped_df$Day)[c(2,6,7,5,1,3,4)])
      for (i in 1:dim(ped_loc)[1])
      {
      ped_df$Lat[ped_df$Sensor_ID == ped_loc$`Sensor ID`[i]] <- ped_loc$Latitude[i]
      ped_df$Lon[ped_df$Sensor_ID == ped_loc$`Sensor ID`[i]] <- ped_loc$Longitude[i]
      }

@@

```{r, eval = FALSE, tidy = TRUE}
missing_glm <- glm(Hourly_Counts ~ Year * HDay * Time + Month)
```

@@

@@

With this project, we want to improve the way we visualise Melbourne's pedestrian activity, and use a predicitive model to get a better understanding of how Melbourne operates

```{r, echo = FALSE}
ggmap(melb) + geom_point(aes(x = Lon, y = Lat, size = Hourly_Counts), alpha = 0.02, data = filter(ped_df, Year == "2016", Month == "March", Time == "9"), colour = "violetred") + scale_size_area(max_size = 10) + guides(colour = guide_legend(override.aes = list(alpha = 1)))
```

>>>>

  melb <- get_map(location=c(mean(range(ped_loc$Longitude)),
                               mean(range(ped_loc$Latitude))),
                                zoom=14)
  melbbw <- get_map(location=c(mean(range(ped_loc$Longitude)),
                               mean(range(ped_loc$Latitude))),
                                zoom=14,
                    color = "bw")
```

@@

```{r, echo = FALSE, message=FALSE, warning=FALSE}
      ped_df$Year <- as.factor(ped_df$Year)
      ped_df$Month <- as.factor(ped_df$Month)
      ped_df$Month <- factor(ped_df$Month, levels(ped_df$Month)[c(5,4,8,1,9,7,6,2,12,11,10,3)])
      ped_df$Time <- as.factor(ped_df$Time)
      ped_df$Timeplot <- as.integer(ped_df$Time)
      ped_df$MTime <- as.integer(ped_df$Time) + ((as.integer(ped_df$Mdate)-1) * 24)
      ped_df$Day <- as.factor(ped_df$Day)
      ped_df$Day <- factor(ped_df$Day, levels(ped_df$Day)[c(2,6,7,5,1,3,4)])

      ped_df$Lat <- 0
      ped_df$Lon <- 0

      for (i in 1:dim(ped_loc)[1])
      {
      ped_df$Lat[ped_df$Sensor_ID == ped_loc$`Sensor ID`[i]] <- ped_loc$Latitude[i]
      ped_df$Lon[ped_df$Sensor_ID == ped_loc$`Sensor ID`[i]] <- ped_loc$Longitude[i]
      }

@@

```{r, eval = FALSE, tidy = TRUE}
missing_glm <- glm(Hourly_Counts ~ Year * HDay * Time + Month,
                   data = ped_df, family = "poisson")
```

@@

---
class: middle, center

## Modelling: evaluating model performance
The criteria we will use to evaluate the model's forecasting performance is the root mean square forecast error:
$$ \text{RMSFE} = \sqrt{\frac{\sum_{t=T+1}^{T+k}(\hat{y_t} - y_t)^2}{k}}$$
where the model is estimated with $T$ observations, and we have $k$ points to forecast (so we need to have $T+k$ observations available)
- evaluating performance of the model on its forecasting accuracy is important, as it will reveal overfitting of data

@@

With this project, we want to improve the way we visualise Melbourne's pedestrian activity
```{r, echo = FALSE, fig.retina = 3, out.height="500px", fig.align = 'center'}
ggmap(melbbw) + geom_point(aes(x = Lon, y = Lat, size = Hourly_Counts), colour = "red", alpha = 0.03, data = filter(ped_df, Year == "2016", Month == "March", Time == "9")) + scale_size_area(max_size = 10) + scale_color_distiller(palette = "PuOr") + guides(colour = guide_legend(override.aes = list(alpha = 1))) + theme_light() + labs(x = NULL, y = NULL)
```

## Visualisation: different methods

```{r, echo = FALSE, message=FALSE, out.height="500px", fig.retina=2}
  ggplot(ped_df[ped_df$Sensor_ID == "3", ]) + 
    geom_smooth(aes(x = Timeplot, y = Hourly_Counts, colour = Month), se = F) + 
    xlab("Time of the Day") + ylab("Pedestrian Count") + 
    ggtitle(paste("Monthly Average Hourly Counts at", ped_loc$`Sensor Description`[ped_loc$`Sensor ID` == "3"], "by Day")) +
    facet_wrap(~ Day, nrow = 2) + theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom") +
    scale_x_continuous(limits = c(0,24))
```
